---
title: "DECIDE score rationale"
author: "Thomas MM; Tom A"
date: "5/18/2021"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(raster)
library(sf)
library(viridis)

source("../../Scripts/modules/filter_distance.R")


```

### Background

Adaptive sampling is the process of optimising the location of new data collection based on the data already gathered. In our case, we want to use the outputs of species distribution models (SDMs) to direct recording to regions that will (hopefully) improve future SDMs. We have decided not to send people to regions that will improve models for a single species. Rather, we want to send people to areas that are likely to improve the models for multiple species. We are doing this by creating a **DECIDE score**, which will be a gridded map at a 100m resolution indicating regions where new data will have the most benefit to future models.

Our current modelling workflow has two main outputs for **each species**. The first is a gridded **probability of presence** map at a 100m resolution, which ranges between 0 and 1. Where the model predicts 0, a species is not predicted to occur; where the model predicts 1, the species is definitely predicted to occur. The second layer is an **uncertainty** or **error** map (the examples below are from old models, so should be taken with a pinch of salt). This is also a gridded map at 100m resolution that ranges between 0 and 1; higher values indicate regions where the model is uncertain in its prediction, and vice versa.


```{r prob_pres, warning = F, message=F, echo = F}

adgr_p <- raster(list.files('../../Data/example_for_rich/rf_models/', pattern = 'meanpred', full.names = T)[5])
adgr_v <- raster(list.files('../../Data/example_for_rich/rf_models/', pattern = 'quantile', full.names = T)[5])

uk_map <- st_as_sf(getData("GADM", country = "GBR", level = 1, path='../../Data/'))
uk_map <- st_transform(uk_map, 27700)
gb_map <- as_Spatial(uk_map[uk_map$NAME_1 != 'Northern Ireland',])

adgr_pc <- mask(adgr_p, gb_map)
adgr_vc <- mask(adgr_v, gb_map)

par(mfrow = c(1,2))
par(mar = c(3,3,3,5))

plot(adgr_pc, main='Broad-bordered White Underwing\nprobability of presence', col=magma(50), cex.main=0.9)
plot(adgr_vc, main='Broad-bordered White Underwing\nuncertainty', col=magma(50), cex.main=0.9)

par(mfrow = c(1,1))


```

To create a DECIDE score across species, we first need to combine the probability of presence and uncertainty maps for each species separately. If we were taking a purely data-driven approach to this problem, we would send people to regions with the greatest uncertainty. This is because new data from areas of high uncertainty are likely to be particularly useful for improving the models. However, many of these places are likely to have very few species present, as some of our other work has shown that model uncertainty is highest in areas with middle probability of presence values (i.e. where probability of presence is around 0.5). A major part of the DECIDE project is about trying to engage people in the process of collecting data to improve our models. Sending people (particularly new recorders) to areas where they are unlikely to see anything, could risk alienating them and affecting their trust in the DECIDE project's nudges. Therefore, we need to create a score that identifies areas where: 

1. the models are the most uncertain and,
2. at least some species are likely to be present.

### Species-level adaptive sampling score

To create a score for each species separately, we have chosen to combine the uncertainty and probability of presence layers by doing:

$$ProbabilityOfPresence  * \sqrt{Uncertainty}$$

Doing this for each species creates a gridded map showing us where new observations will be most beneficial to that species' model. Taking the square root of the uncertainty layer means that this score will be more sensitive to changes in probability of presence than to changes in uncertainty. This will therefore give a higher weighting to areas in which species have a high probability of presence than to places with high uncertainty.  

### The cross-species DECIDE score

Once we've generated an adaptive sampling metric for each species, we then want to create the **DECIDE score** that will enable us to direct people where to sample. We have decided to combine the adaptive sampling score for each species by taking their mean, weighted by the probability of presence for each species. This means that species with a very low probability of presence in a grid cell won’t contribute much to the overall DECIDE score; species that are likely to be present in a grid cell will contribute more. This is important because it means that the DECIDE score won't be influenced by species that aren't likely to be present in an area. For example, species that are restricted to the Scottish Highlands will have no (well actually, extremely little) effect on the DECIDE scores for areas around Wallingford. The figure below shows the GB-wide DECIDE scores for moths and butterflies.

```{r GB_decide, echo = F, warning=FALSE,message=FALSE, fig.width = 8}

bgb <- raster('../../Data/species_data/decide_scores/butterfly_weighted_prob_pres_GB_decide_score.grd')
mgb <- raster('../../Data/species_data/decide_scores/moth_weighted_prob_pres_GB_decide_score.grd')

par(mfrow = c(1,2))
par(mar = c(3,3,3,5))

plot(bgb, main='Butterfly DECIDE score', col=viridis(50))
plot(mgb, main='Moth DECIDE score', col=viridis(50))

par(mfrow = c(1,1))

```

### Other points to mention

It is important to note that there are various aspects of the DECIDE score that can be altered.

1. This DECIDE score is based solely on the models, not user preferences and the other factors that we have previously discussed 
2. Currently, the DECIDE score is completely static for this first iteration. In the final version of the app species will be toggled on and off depending on various filters. For example, species that aren't flying at the current time of year will be removed, species that are of conservation concern could be highlighted, people might be able to choose 'preferred' species etc.
3. The colour scheme is still a work in progress, and can be altered to highlight the best areas.


```{r exit, echo = F}

knitr::knit_exit()

```

## Different DECIDE scores

I want to investigate different ways of combining the probability of presence and uncertainty layers for each species, to see what effect it has on values at a GB and local scale. The three different ways are the current way $$ProbabilityOfPresence * \sqrt{Uncertainty}$$, an equal weighting of presence and uncertainty $$ProbabilityOfPresence*Uncertainty$$ and weighted towards uncertainty $$\sqrt{ProbabilityOfPresence}*Uncertainty$$. The average across all species is then taken, weighted by a species' probability of presence in a grid cell.


```{r comp_dec_scores, message=F, warning=F, echo=F}

sqvar_b <- raster('/data-s3/thoval/sdm_outputs/decide_scores/butterfly_PA_thinned_10000nAbs_decide_score_sqroot_var_preds_weight_mean.grd')
eq_b <- raster('/data-s3/thoval/sdm_outputs/decide_scores/butterfly_PA_thinned_10000nAbs_decide_score_equal_weighting_weight_mean.grd')
sqpred_b <- raster('/data-s3/thoval/sdm_outputs/decide_scores/butterfly_PA_thinned_10000nAbs_decide_score_sqroot_preds_var_weight_mean.grd')


sqvar_m <- raster('/data-s3/thoval/sdm_outputs/decide_scores/moth_PA_thinned_10000nAbs_decide_score_sqroot_var_preds_weight_mean.grd')
eq_m <- raster('/data-s3/thoval/sdm_outputs/decide_scores/moth_PA_thinned_10000nAbs_decide_score_equal_weighting_weight_mean.grd')
sqpred_m <- raster('/data-s3/thoval/sdm_outputs/decide_scores/moth_PA_thinned_10000nAbs_decide_score_sqroot_preds_var_weight_mean.grd')


par(mfrow = c(1,3))
plot(sqvar_b, col=viridis(50), main = 'butterfly, sqroot var')
plot(eq_b, col=viridis(50), main = 'equal weighting')
plot(sqpred_b, col=viridis(50), main = 'sqroot pred')


plot(sqvar_m, col=viridis(50), main = 'moth, sqroot var')
plot(eq_m, col=viridis(50), main = 'equal weighting')
plot(sqpred_m, col=viridis(50), main = 'sqroot pred')
par(mfrow = c(1,1))



```

Going to zoom in on Aberlady Bay, Scotland.

```{r aberlady, message = F, warning = F, echo=F}

location <- c(-2.860564, 56.014902)
distance <- 5000

r_list <- list(sqvar_b, eq_b, sqpred_b, sqvar_m, eq_m, sqpred_m)
p_out <- list()

for(i in 1:length(r_list)){
  
  fc <- filter_distance(r_list[[i]],
                        location = location, 
                        distance = distance,
                        method = 'buffer')
  p_out[[i]] <- fc  
  
}

par(mfrow = c(2,3))
plot(p_out[[1]], col=viridis(50), main = 'butterfly, sqroot var')
plot(p_out[[2]], col=viridis(50), main = 'equal weighting')
plot(p_out[[3]], col=viridis(50), main = 'sqroot pred')


plot(p_out[[4]], col=viridis(50), main = 'moth, sqroot var')
plot(p_out[[5]], col=viridis(50), main = 'equal weighting')
plot(p_out[[6]], col=viridis(50), main = 'sqroot pred')
par(mfrow = c(1,1))


```

As you can see, the methods of combining DECIDE scores across species show some differences at the GB scale, but look very similar at the local scale. I guess this probably has something to do with the relationship between probability of presence and uncertainty in any given region - i.e. that uncertainty is usually higher at intermediate values of probability of presence. Need to investigate different ways of combining across species to see what effect that has on the DECIDE Score.

## What is the DECIDE score telling us?

### Exploring correlations

```{r setup_corrs}


sum_pres <- sum(cellStats(preds))


```
