---
title: "DECIDE score plots"
author: "Thomas MM"
date: "6/22/2021"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = F, message = F)

library(tidyverse)
library(doParallel)
library(foreach)
library(raster)
library(viridis)
library(scales)
library(ggmap)
library(osmdata)

source("../../Scripts/modules/filter_distance.R")

```


```{r species_data, warning=F, echo = F, message = F, include = F}

# model = c('rf', 'lr', 'gam')
taxa = 'moth'
pseudoabs = 'PA_thinned_10000nAbs'


model_locs <- paste0('/data-s3/thoval/sdm_outputs/', taxa, '/combined_model_outputs/', pseudoabs)

names <- gsub(pattern = '_PA_thinned_10000nAbs_weightedmeanensemble.grd', replacement = '', 
              list.files(model_locs, 
                         pattern='_weightedmeanensemble.grd'))
# names

# sdm outputs for each species
species_stack <- list()

# error outputs
error_out <- list()

for(i in 1:length(names)){
  
  print(names[i])
  
  # initiate model list within for loop so that it gets replaced when starting a new species
  # otherwise we might get some weird overlaps
  
  # mean predictions
  mp <- list.files(model_locs, 
                   pattern = paste0(names[i], "_", pseudoabs, "_weightedmeanensemble.grd"),
                   full.names = TRUE)
  
  mod_preds <- raster::stack(mp)
  names(mod_preds) <- paste0(names[i], '_mean_pred')
  
  
  
  # quantile range
  qr <- list.files(model_locs, 
                   pattern = paste0(names[i], "_", pseudoabs, "_rangeensemblequantiles.grd"),
                   full.names = TRUE)
  
  qrnge <- raster::stack(qr)
  names(qrnge) <- paste0(names[i], '_quantile_range')
  
  species_stack[[i]] <- raster::stack(mod_preds, qrnge)
  
}


```


```{r loc_dist}

# location = c(-0.385843, 51.289680) # bookham common
# location = c(-2.860564, 56.014902) # aberlady
# location = c(-1.503560, 54.141389) # sharow
# location = c(-2.247979, 50.632404) # lulworth cove
# location = c(-1.461817, 54.440435) # great smeaton
# location = c(-1.110557, 51.602436) # wallingford


distance = 5000

```


```{r crop, warning = F, include = F}

registerDoParallel(7)

# out_cropped <- list()
system.time(
  out_cropped <- foreach(s = 1:length(species_stack)) %dopar% {
    
    print(s)
    
    sp <- species_stack[[s]]
    
    
    # crop the prediction
    crop_pred <- filter_distance(obj = subset(sp, grep(pattern = 'mean_pred',
                                                       names(sp))),
                                 method = 'buffer',
                                 distance = distance,
                                 location = location)
    
    # crop the error
    crop_err <- filter_distance(obj = subset(sp, grep(pattern = 'quantile_range',
                                                      names(sp))),
                                method = 'buffer',
                                distance = distance,
                                location = location)
    
    if(length(names(crop_pred))>1){
      # get the mean
      m_pred_av <- calc(crop_pred,
                        mean, na.rm = T)
      names(m_pred_av) <- 'predictions'
      
      
      m_quant_av <- calc(crop_err,
                         mean, na.rm = T)
      names(m_quant_av) <- 'error'
    } else {
      
      m_pred_av <- crop_pred
      m_quant_av <- crop_err
      
    }
    
    out_rasts <- list(m_pred_av, m_quant_av)
    names(out_rasts) <- c('predictions', 'quantile_var')
    
    return(out_rasts)
    
  }
)

registerDoSEQ()

names(out_cropped) <- names


# get the cropped probability of presence
preds <- stack(lapply(1:length(out_cropped), FUN = function(x) out_cropped[[x]]$predictions))
names(preds) <- names(out_cropped)

# get the cropped variation
var <- stack(lapply(1:length(out_cropped), FUN = function(x) out_cropped[[x]]$quantile_var))
names(var) <- names(out_cropped)


```


```{r, message = F, warning = F}

register_google(key = "AIzaSyDzfFo6Jq1g65BauVhLL-SexOYS5OuH-kA")

gmap <- get_map(location = location, zoom = 12)

## original decide
ogd <- projectRaster(raster::weighted.mean(sqrt(var) * (preds), preds), crs="+proj=longlat +datum=WGS84") %>% as.data.frame(xy=T)

ogd_p <- ggmap(gmap) +
  geom_tile(data = ogd, aes(x=x,y=y, fill = layer), alpha = 0.7) +
  coord_quickmap() +
  scale_fill_gradient(high = 'red', low = '#ffffcc', na.value = NA, name = 'DECIDE score') +
  labs(x= '', y = '')


## variation only
vo <- projectRaster(mean(var), crs="+proj=longlat +datum=WGS84") %>% as.data.frame(xy=T)

vo_p <- ggmap(gmap) +
  geom_tile(data = vo, aes(x=x,y=y, fill = layer), alpha = 0.7) +
  coord_quickmap() +
  scale_fill_gradient(high = 'red', low = '#ffffcc', na.value = NA, name = 'DECIDE score') +
  labs(x= '', y = '')


## equal weighting
eq <- mean((var)*(preds))

# project raster to lat long for use with ggmap
eqp <- projectRaster(eq, crs="+proj=longlat +datum=WGS84") %>% as.data.frame(xy=T)

eqp_p <- ggmap(gmap) +
  geom_tile(data = eqp, aes(x=x,y=y, fill = layer), alpha = 0.7) +
  coord_quickmap() +
  scale_fill_gradient(high = 'red', low = '#ffffcc', na.value = NA, name = 'DECIDE score') +
  labs(x= '', y = '')


## variation upweighted
vp <- projectRaster(mean(sqrt(var)*(preds)), crs="+proj=longlat +datum=WGS84") %>% as.data.frame(xy=T)

vp_p <- ggmap(gmap) +
  geom_tile(data = vp, aes(x=x,y=y, fill = layer), alpha = 0.7) +
  coord_quickmap() +
  scale_fill_gradient(high = 'red', low = '#ffffcc', na.value = NA, name = 'DECIDE score') +
  labs(x= '', y = '')


## Cells with the highest decide scores
q90 <- projectRaster(calc(var, fun=function(x) quantile(x, .9, na.rm=TRUE)), 
                    crs="+proj=longlat +datum=WGS84") %>% as.data.frame(xy=T)

q90_p <- ggmap(gmap) +
  geom_tile(data = q90, aes(x=x,y=y, fill = layer), alpha = 0.7) +
  coord_quickmap() +
  scale_fill_gradient(high = 'red', low = '#ffffcc', na.value = NA, name = 'DECIDE score') +
  labs(x= '', y = '')




```


Below are three different ways of calculating the DECIDE score. The first plot is the average uncertainty of all species in a grid cell; meaning that all species are weighted equally to get the final value. The second plot is also based on uncertainty only, but only takes into account the most uncertain species when calculating the final value. Finally, the third plot calculates the DECIDE score by accounting for uncertainty and the probability of presence.


```{r plots, fig.width=8}


# p1 <- ogd_p + ggtitle('Original DECIDE score')
p2 <- vo_p  + ggtitle('Average uncertainty')
p3 <- eqp_p + ggtitle('Presence and uncertainty equally weighted')
# p4 <- vp_p  + ggtitle('Upweighted uncertainty with presence')
p5 <- q90_p + ggtitle('Highest uncertainty')

# p1
p2 
p5
p3
# p4

```


```{r}
# 
# spdf <- read.csv('../../Data/species_data/moth/DayFlyingMoths_East_Norths.csv')
# 
# p2 + geom_point(data = spdf, aes(x=lon,y=lat), colour = 'black', size = 0.8, pch = 20)

```




