---
title: "Providing DECIDE nudges"
author: "Thomas MM; Tom A"
date: "6/1/2021"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(raster)
library(patchwork)


source("../../Scripts/modules/filter_distance.R")
source("../../Scripts/modules/convert_raster.R")
source("../../Scripts/modules/nudge_thin.R")
source("../../Scripts/modules/nudge_select.R")
source("../../Scripts/modules/nudge_list.R")
source("../../Scripts/modules/nudge_accessible.R")
source("../../Scripts/modules/nudge_metadata.R")
source("../../Scripts/modules/load_gridnumbers.R")


```


This document will run through a potential workflow for translating the DECIDE score into nudges. The nudging process essentially follows these steps:

1. Generate a list of potential nudge locations
2. Select a user-determined subset of these nudge locations
3. Determine which nudge locations are near to accessible areas
4. Thin the nudges based on a user-defined grid

First, need to prepare the raster data for use; load it in and crop to a location. In this case, the location is 5km around Wallingford.

```{r dec_rast, message = FALSE, warning = FALSE}

mgb <- raster('../../Data/species_data/decide_scores/moth_weighted_prob_pres_GB_decide_score.grd')


location = c(-1.110557, 51.602436) # wallingford
# location = c(-2.730696, 54.026759) # quernmore
distance = 5000

# crop
wall_m <- filter_distance(mgb,
                          location=location,
                          distance=distance,
                          method = 'buffer')

```

### 1. List of potential nudges

The **'nudge_list'** function generates a list of potential nudge locations . This takes the raster we created above and converts it to a data frame. We can provide the function with a quantile value and a cutoff 'TRUE/FALSE' statement. The function identifies the values that exceed the quantile for the area of interest (defined by 'cutoff_value') and, if cutoff = TRUE, it will remove all the points falling below it. If cutoff = FALSE, it retains all the values irrespective of their score. The function returns a data frame of potential nudge locations whose length is determined by 'prop'; its value defines the proportion of the total number of cells in an area that we want returned. The values can be weighted by the DECIDE score value, so that locations with high decide scores are preferentially selected. These values are often low, so I have inflated the effect of the DECIDE score by providing an inflation factor ('weight_inflation').

Below are some examples of the effects of the different parameters on the nudges returned. In the cases below, I have shown all combinations of weightings being turned on and off and the cutoff being turned on and off. NB if cutoff = FALSE and weight = FALSE, the function will return a random number (determined by 'prop') of background points.

```{r nudge_list_df_plots, echo = TRUE}

nud_prop01_cut_F <-  nudge_list(wall_m,
                                prop = 0.01, 
                                cutoff_value = 0.9, 
                                cutoff = FALSE,
                                weight = TRUE,
                                weighting_column = 'weights',
                                weight_inflation = 50)


nud_prop01_cut_F_unweight <-  nudge_list(wall_m,
                                         prop = 0.01, 
                                         cutoff_value = 0.9, 
                                         cutoff = FALSE,
                                         weight = FALSE,
                                         weighting_column = 'weights',
                                         weight_inflation = 50)

nud_prop01_cut_T <-  nudge_list(wall_m,
                                prop = 0.01, 
                                cutoff_value = 0.9, 
                                cutoff = TRUE,
                                weight = TRUE,
                                weighting_column = 'weights',
                                weight_inflation = 50)


nud_prop01_cut_T_unweight <-  nudge_list(wall_m,
                                         prop = 0.01, 
                                         cutoff_value = 0.9, 
                                         cutoff = TRUE,
                                         weight = FALSE,
                                         weighting_column = 'weights',
                                         weight_inflation = 50)


```


```{r nudge_list_plots, echo = FALSE, message = FALSE, warning = FALSE, fig.width=12}

nl_p1 <- ggplot() +
  geom_sf(data = conv_rast(wall_m, 27700), aes(fill = layer), col = NA) +
  geom_point(data = nud_prop01_cut_F, aes(x=lon,y=lat, col='Weighted')) +
  geom_point(data = nud_prop01_cut_F_unweight, aes(x=lon,y=lat, col='Unweighted')) +
  
  scale_fill_continuous(type= 'viridis', name = 'DECIDE score') +
  scale_colour_manual(name = '', values = c('black', 'red')) +
  theme_bw() +
  labs(x='', y='', title = 'No cutoff')

nl_p2 <- ggplot() +
  geom_sf(data = conv_rast(wall_m, 27700), aes(fill = layer), col = NA) +
  geom_point(data = nud_prop01_cut_T, aes(x=lon,y=lat, col='Weighted')) +
  geom_point(data = nud_prop01_cut_T_unweight, aes(x=lon,y=lat, col='Unweighted')) +
  
  scale_fill_continuous(type= 'viridis', name = 'DECIDE score') +
  scale_colour_manual(name = '', values = c('black', 'red')) +
  theme_bw() +
  labs(x='', y='', title = 'With cutoff')

nl_p1 + nl_p2 + plot_layout(guides = 'collect')

```

Here is a new data frame of weighted nudges from above and below the 90th percentile to use with the later functions.

```{r nudge_list, echo=TRUE, message = FALSE, warning = FALSE}

# get a list of potential nudge locations
nudges <- nudge_list(wall_m,
                     n=500,
                     # prop = 0.01,
                     cutoff_value = 0.9, 
                     cutoff = FALSE,
                     weight = TRUE,
                     weighting_column = 'decide score',
                     weight_inflation = 10)
head(nudges)


```


```{r nudge_list_plot, warning = FALSE, message = FALSE, echo = FALSE}

nl_p_nudge1 <- ggplot() +
  geom_sf(data = conv_rast(wall_m, 27700), aes(fill = layer), col = NA) +
  geom_point(data = nudges, aes(x=lon,y=lat, col='Weighted nudges'))  +
  
  scale_fill_continuous(type= 'viridis', name = 'DECIDE score') +
  scale_colour_manual(name = '', values = 'black') +
  theme_bw() +
  labs(x='', y='', title = 'All nudges') +
  theme(legend.margin=margin(t=-0.25,l=0.05,b=0.0,r=0.05, unit='cm')) 
nl_p_nudge1

```



### 2. Select a subset of these nudges

We can get a specific number of these nudges using the function 'nudge_select'. This takes the total number of nudges we want (determined by 'n') and samples the data frame outputted from the 'nudge_list' function. The function 'nudge_select' can also be weighted by the DECIDE score, which can be inflated by a user-determined amount, to return only the highest value locations. The reason this function is split from the last one is that I can imagine a user wanting to randomise the nudges they are sent. With the functions split, it means that we would only have to run 'nudge_select' as opposed to also running the 'nudge_list' function above when users want a new sample of nudges. Currently, the areas being shown are only within 5km of the user. However, 'nudge_list' is very likely to be slow with larger rasters because of the 'raster::as.data.frame' function (although I haven't directly tested this yet but have encountered problems in the past). The 'all nudges' panel in the plots below is the same as the previous plot and is being compared to the 100 nudges selected from it weighted by the DECIDE score.

This function isn't needed anymore after changes to the function above.


```{r nudge_subset}

# nudge_subset <- nudge_select(nudge_df = nudges, 
#                              n = 100,
#                              cutoff = FALSE,
#                              weight = TRUE,
#                              weighting_column = 'dec_score',
#                              weight_inflation = 10)


```

```{r nudge_subset_plot, echo = FALSE, warning=FALSE, fig.width=11}


# nl_p_nudge2 <- ggplot() +
#   geom_sf(data = conv_rast(wall_m, 27700), aes(fill = layer), col = NA) +
#   geom_point(data = nudge_subset, aes(x=lon,y=lat, col='Weighted nudges subset'))  +
#   
#   scale_fill_continuous(type= 'viridis', name = 'DECIDE score') +
#   scale_colour_manual(name = '', values = 'red') +
#   theme_bw() +
#   labs(x='', y='', title = 'Subsetted nudges') + 
#   theme(legend.margin=margin(t=-0.25,l=0.05,b=0.0,r=0.05, unit='cm')) 
# 
# nl_p_nudge1 + nl_p_nudge2 + plot_layout(guides = 'collect')

```



### 3. Find nudges close to accessible areas

We need to find the nudges that are actually accessible to recorders. For this, we can use the 'nudge_accessible' function. It takes an sf dataframe or just normal dataframe of nudges,  accessible areas (i.e. footpaths etc.) and a buffer distance, and returns all the nudges that fall within this region. The plot output is optional but is a nice 'sense check'.

```{r load_access_locs, echo = F, message = F, warning = F, results='hide'}

## function to choose points close to accessible features 

# first load the accessible areas

# load a UK grid - currently 10km
grid <- st_read('/data/notebooks/rstudio-setupconsthomas/DECIDE_constraintlayers/Data/raw_data/UK_grids/uk_grid_10km.shp')
st_crs(grid) <- 27700

# find grid numbers that location + buffered zone covers
grid_numbers <- load_gridnumbers(location = location,
                                 distance = distance,
                                 grid = grid)

# shape data
# point to shape locations - move data to a common directory
{# england
  prow_loc <- ("/data/notebooks/rstudio-setupconsthomas/DECIDE_constraintlayers/Data/raw_data/rowmaps_footpathbridleway/rowmaps_footpathbridleway/gridded_data_10km/")
  grnspc_loc <- "/data/notebooks/rstudio-setupconsthomas/DECIDE_constraintlayers/Data/raw_data/OS_greenspaces/OS Open Greenspace (ESRI Shape File) GB/data/gridded_greenspace_data_10km/"
  accspnt_loc <- "/data/notebooks/rstudio-setupconsthomas/DECIDE_constraintlayers/Data/raw_data/OS_greenspaces/OS Open Greenspace (ESRI Shape File) GB/data/gridded_accesspoint_data_10km/"
  access_land_loc <- "/data/notebooks/rstudio-setupconsthomas/DECIDE_constraintlayers/Data/raw_data/CRoW_Act_2000_-_Access_Layer_(England)-shp/gridded_data_10km/"
  nat_trust_loc <- '/data/notebooks/rstudio-setupconsthomas/DECIDE_constraintlayers/Data/raw_data/national_trust/gridded_data_10km/'
  lond_path_loc <- '/data/notebooks/rstudio-setupconsthomas/DECIDE_constraintlayers/Data/raw_data/greater-london-latest-free/london_gridded_data_10km/'
  
  
  # scotland
  core_paths <- "/data/notebooks/rstudio-setupconsthomas/DECIDE_constraintlayers/Data/raw_data/Scotland/core_paths/gridded_data_10km/"
  pub_acc <- "/data/notebooks/rstudio-setupconsthomas/DECIDE_constraintlayers/Data/raw_data/Scotland/public_access_wiat/gridded_data_10km/"
  cairn <- "/data/notebooks/rstudio-setupconsthomas/DECIDE_constraintlayers/Data/raw_data/Scotland/cairngorms/gridded_data_10km/"
  tross <- "/data/notebooks/rstudio-setupconsthomas/DECIDE_constraintlayers/Data/raw_data/Scotland/lochlomond_tross/gridded_data_10km/"
  cons_sites <- "/data/notebooks/rstudio-setupconsthomas/DECIDE_constraintlayers/Data/raw_data/Scotland/local_nature_conservation_sites/gridded_data_10km/"
  nat_sites <- "/data/notebooks/rstudio-setupconsthomas/DECIDE_constraintlayers/Data/raw_data/Scotland/local_nature_reserves/gridded_data_10km/"
  wild <- "/data/notebooks/rstudio-setupconsthomas/DECIDE_constraintlayers/Data/raw_data/Scotland/wildland_scotland/gridded_data_10km/"
}

# extract shapes for location of interest
system.time(
  acc_loc <- lapply(c(1:length(grid_numbers)), FUN = function(n){
    
    
    ####    ENGLAND    ####
    # prow .shp
    prow_files <- list.files(prow_loc,
                             full.names = T,
                             pattern = paste0('_', grid_numbers[n], '.shp'))
    
    if(length(prow_files) != 0) {
      prow <- sf::st_read(prow_files, quiet = TRUE)
      st_crs(prow) <- 27700
    } else { prow <- NULL }
    
    # greenspaces .shp
    grnspc_files <- list.files(grnspc_loc,
                               full.names = T,
                               pattern = paste0('_', grid_numbers[n], '.shp'))
    
    if(length(grnspc_files) != 0){
      grnspc <- sf::st_read(grnspc_files, quiet = TRUE)
      st_crs(grnspc) <- 27700
    } else { grnspc <- NULL }
    
    # access points .shp
    accs_files <- list.files(accspnt_loc,
                             full.names = T,
                             pattern = paste0('_', grid_numbers[n], '.shp'))
    
    if(length(accs_files) != 0){
      accspnt <- sf::st_read(accs_files, quiet = TRUE)
      st_crs(accspnt) <- 27700
    } else{ accspnt <- NULL }
    
    # access land .shp
    accslnd_files <- list.files(access_land_loc,
                                full.names = T,
                                pattern = paste0('_', grid_numbers[n], '.shp'))
    
    if(length(accslnd_files) != 0){
      accslnd <- sf::st_read(accslnd_files, quiet = TRUE)
      st_crs(accslnd) <- 27700
    } else { accslnd <- NULL }
    
    # national trust .shp
    national_trust_files <- list.files(nat_trust_loc,
                                       full.names = T,
                                       pattern = paste0('_', grid_numbers[n], '.shp'))
    
    if(length(national_trust_files) != 0){
      national_trust <- sf::st_read(national_trust_files, quiet = TRUE)
      st_crs(national_trust) <- 27700
    } else { national_trust <- NULL }
    
    # london shapes .shp
    london_files <- list.files(lond_path_loc,
                               full.names = T,
                               pattern = paste0('_', grid_numbers[n], '.shp'))
    
    if(length(london_files) != 0){
      london_paths <- sf::st_read(london_files, quiet = TRUE)
      st_crs(london_paths) <- 27700
    } else { london_paths <- NULL }
    
    
    ####    SCOTLAND    ####
    
    # core paths
    cp_files <- list.files(core_paths,
                           full.names = T,
                           pattern = paste0('_', grid_numbers[n], '.shp'))
    
    if(length(cp_files) != 0){
      c_paths <- sf::st_read(cp_files, quiet = TRUE)
      st_crs(c_paths) <- 27700
    } else { c_paths <- NULL }
    
    
    # public access
    # paths
    pacc_path_files <- list.files(pub_acc,
                                  full.names = T,
                                  pattern = paste0('paths_gridnumber_', grid_numbers[n], '.shp'))
    
    if(length(pacc_path_files) != 0){
      pacc_paths <- sf::st_read(pacc_path_files, quiet = TRUE)
      st_crs(pacc_paths) <- 27700
    } else { pacc_paths <- NULL }
    
    # woods
    pacc_wood_files <- list.files(pub_acc,
                                  full.names = T,
                                  pattern = paste0('wood_gridnumber_', grid_numbers[n], '.shp'))
    
    if(length(pacc_wood_files) != 0){
      pacc_wood <- sf::st_read(pacc_wood_files, quiet = TRUE)
      st_crs(pacc_wood) <- 27700
    } else { pacc_wood <- NULL }
    
    
    ## national parks
    # cairngorm
    cairn_files <- list.files(cairn,
                              full.names = T,
                              pattern = paste0('_', grid_numbers[n], '.shp'))
    
    if(length(cairn_files) != 0){
      cairn_shp <- sf::st_read(cairn_files, quiet = TRUE)
      st_crs(cairn_shp) <- 27700
    } else { cairn_shp <- NULL }
    
    
    # trossacks
    tross_files <- list.files(tross,
                              full.names = T,
                              pattern = paste0('_', grid_numbers[n], '.shp'))
    
    if(length(tross_files) != 0){
      tross_shp <- sf::st_read(tross_files, quiet = TRUE)
      st_crs(tross_shp) <- 27700
    } else { tross_shp <- NULL }
    
    
    # conservation sites
    cons_files <- list.files(cons_sites,
                             full.names = T,
                             pattern = paste0('_', grid_numbers[n], '.shp'))
    
    if(length(cons_files) != 0){
      cons_shp <- sf::st_read(cons_files, quiet = TRUE)
      st_crs(cons_shp) <- 27700
    } else { cons_shp <- NULL }
    
    
    # natural conservation sites
    nat_files <- list.files(nat_sites,
                            full.names = T,
                            pattern = paste0('_', grid_numbers[n], '.shp'))
    
    if(length(nat_files) != 0){
      nat_shp <- sf::st_read(nat_files, quiet = TRUE)
      st_crs(nat_shp) <- 27700
    } else { nat_shp <- NULL }
    
    
    # wildland sites
    wild_files <- list.files(wild,
                             full.names = T,
                             pattern = paste0('_', grid_numbers[n], '.shp'))
    
    if(length(wild_files) != 0){
      wild_shp <- sf::st_read(wild_files, quiet = TRUE)
      st_crs(wild_shp) <- 27700
    } else { wild_shp <- NULL }
    
    return(list(prow, grnspc, accspnt, accslnd, national_trust, london_paths, ## england
                c_paths, pacc_paths, pacc_wood, cairn_shp, tross_shp, cons_shp, nat_shp, wild_shp)) ## scotland
    
  })
)

all_outs <- do.call(Map, c(rbind, acc_loc))

# crop shapes to exact region of interest
final_acc_loc <- lapply(all_outs, filter_distance,
                        location = location,
                        distance = distance,
                        method = 'buffer')



```


```{r nudge_accessible}


accessible_nudges <- nudge_accessible(nudges_df = nudges,
                                      access_layers = final_acc_loc,
                                      buffer = 200,
                                      crs = 27700,
                                      lon = 'lon',
                                      lat = 'lat',
                                      plot = TRUE)


```

### 4. Thin the nudges

There is unlikely to be much point in providing the user with nudges that are very close together. The function 'nudge_thin' creates a grid whose resolution is defined by the user and samples from the points within each of the grid cells. The user also states how many points to sample from each cell and can specify whether to use a square or hexagonal grid (the latter being better for large areas). This uses the original raster layer and the 'nudge_select' function (although will also work with the 'nudge_list' function) and provides as output: a data frame with the thinned nudges, the grid as an sf object and the plot. The plot takes a while to produce (a few seconds), but obviously won't need to be produced for the app. This method of 'thinning' the nudges does mean that some nudges will be close together if they are in adjacent cells. I don't think this matters too much and I couldn't figure out how to do it based on raw distances between each point.


```{r thin_nudge, warning = FALSE, message = FALSE, echo = T}


thinned_nudges <- nudge_thin(decide_raster = wall_m,
                             nudge_df = accessible_nudges$nudges,
                             lon = 'lon',
                             lat = 'lat',
                             crs = 27700, 
                             buffer_distance = 1000,
                             sample_num = 1)

thinned_nudges$plot

```

### "Final" nudges

The nudges can then be plotted for a user!

```{r final_nudges, echo = FALSE, message = FALSE, warning = FALSE}

ggplot() +
  geom_sf(data = conv_rast(wall_m, 27700), aes(fill=layer), colour = NA) +
  geom_sf(data = thinned_nudges$nudges, aes(colour = 'Suggested sampling\nlocations')) +
  scale_fill_continuous(type = 'viridis', name = 'DECIDE score') +
  scale_colour_manual(name = '', values = 'red') +
  theme_bw()


```

## Get metadata for nudges

If they then click on a point they could get some metadata about the layers that the point is close to. For now, this is only information from the accessible layers, but it would be easy to expand this to what landcover type the nudge is in, for example.

```{r meta, warning=F}

metadata <- nudge_metadata(nudges_df = thinned_nudges$nudges,
                           access = final_acc_loc,
                           buffer = 200)
metadata

metadata_single <- nudge_metadata(nudges_df = thinned_nudges$nudges[3,],
                                  access = final_acc_loc,
                                  buffer = 200)
metadata_single

```

### Timings

The nudging process doesn't take that long but here are some time estimates for the same process without any of the plots being generated (as it would be in the app). I've run the nudge_metadata code for all the points and a single point on its own to understand the time difference between them.


```{r timings, warning = FALSE, message = FALSE}

t_nl <- (system.time(nudges <- nudge_list(wall_m,
                                          n=100,
                                          # prop = 0.1, 
                                          cutoff_value = 0.9, 
                                          cutoff = FALSE,
                                          weight = TRUE,
                                          weight_inflation = 10)))

# t_ns <- system.time(nudge_subset <- nudge_select(nudge_df = nudges, 
#                                                  n = 100,
#                                                  cutoff = FALSE,
#                                                  weight = TRUE,
#                                                  weighting_column = 'dec_score',
#                                                  weight_inflation = 10))

t_na <- system.time(accessible_nudges <- nudge_accessible(nudges_df = nudges,
                                                          access_layers = final_acc_loc,
                                                          buffer = 200,
                                                          crs = 27700,
                                                          lon = 'lon',
                                                          lat = 'lat',
                                                          plot = FALSE))

t_nt <- system.time(thinned_nudges <- nudge_thin(decide_raster = wall_m,
                                                 nudge_df = accessible_nudges$nudges,
                                                 lon = 'lon',
                                                 lat = 'lat',
                                                 crs = 27700, 
                                                 buffer_distance = 1000,
                                                 sample_num = 1))

t_md <- system.time(metadata <- nudge_metadata(nudges_df = thinned_nudges$nudges,
                                               access = final_acc_loc,
                                               buffer = 200))

t_mds <- system.time(metadata_single <- nudge_metadata(nudges_df = thinned_nudges$nudges[1,],
                                                       access = final_acc_loc,
                                                       buffer = 200))



time_df <- rbind(t_nl, #t_ns, 
                 t_na, t_nt, t_md, t_mds) %>% 
  as.data.frame() 

rownames(time_df) <- c('nudge_list', #'nudge_select', 
'nudge_accessible', 'nudge_thin', 'nudge_metadata_all', 'nudge_metadata_single')

time_df <- time_df %>% 
  rownames_to_column(var = 'Function')

time_df


```

The total elapsed time for nudging and getting a single point's metadata is `r sum(time_df$elapsed[c(1:4,6)])`. 

