---
title: "DECIDE Score"
author: "Thomas MM; Tom A"
date: "4/8/2021"
output: html_document
---

```{r setup, include=FALSE, echo = FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(spatstat)



```

## Load species data

```{r species_data}

model = c('rf', 'lr', 'gam')
taxa = 'moth'

# get a list of all the species that appear in the outputs
spp_names_lr <- unique(gsub(pattern="lr_SDMs_|_meanpred.grd|_quantilemaxmin.grd|_quantilerange.grd", replacement = '', 
                            x = list.files(paste0('/data-s3/thoval/sdm_outputs/', taxa, '/lr'), pattern = '.grd')))

spp_names_rf <- unique(gsub(pattern="rf_SDMs_|_meanpred.grd|_quantilemaxmin.grd|_quantilerange.grd", replacement = '', 
                            x = list.files(paste0('/data-s3/thoval/sdm_outputs/', taxa, '/rf'), pattern = '.grd')))

spp_names_gam <- unique(gsub(pattern="gam_SDMs_|_meanpred.grd|_quantilemaxmin.grd|_quantilerange.grd", replacement = '', 
                             x = list.files(paste0('/data-s3/thoval/sdm_outputs/', taxa, '/gam'), pattern = '.grd')))

names <- unique(c(spp_names_lr, spp_names_rf, spp_names_gam))

# sdm outputs for each species
species_stack <- list()

# error outputs
error_out <- list()

for(i in 1:length(names)){
  
  print(names[i])
  
  # initiate model list within for loop so that it gets replaced when starting a new species
  # otherwise we might get some weird overlaps
  model_stack <- list()
  errored_models <- list()
  
  for(m in 1:length(model)){
    
    check_models <- list.files(paste0('/data-s3/thoval/sdm_outputs/', taxa, '/', model[m]), 
                               pattern = paste0(names[i]),
                               full.names = TRUE)
    
    if(length(check_models)<=1){
      
      print(paste('!!!   model', model[m], 'failed for species', names[i], '  !!!'))
      
      errored_models[[m]] <- data.frame(taxa = taxa, 
                                        species = names[i], 
                                        model = model[m])
      
      next
    }
    
    # mean predictions
    mp <- list.files(paste0('/data-s3/thoval/sdm_outputs/', taxa, '/', model[m]), 
                     pattern = paste0(names[i], "_meanpred.grd"),
                     full.names = TRUE)
    
    mod_preds <- raster::stack(mp)
    names(mod_preds) <- paste0(names[i], '_', model[m],'_mean_pred')
    
    
    
    # quantile min/max
    mm <- list.files(paste0('/data-s3/thoval/sdm_outputs/', taxa, '/', model[m]), 
                     pattern = paste0(names[i], "_quantilemaxmin.grd"),
                     full.names = TRUE)
    
    qminmax <- raster::stack(mm)
    names(qminmax) <- c(paste0(names[i], '_', model[m],'_min'), paste0(names[i], '_', model[m],'_max'))
    
    
    # quantile range
    qr <- list.files(paste0('/data-s3/thoval/sdm_outputs/', taxa, '/', model[m]), 
                     pattern = paste0(names[i], "_quantilerange.grd"),
                     full.names = TRUE)
    
    qrange <- raster::stack(qr)
    names(qrange) <- paste0(names[i], '_', model[m], '_quantile_range')
    
    
    # stack all from one model together
    model_stack[[m]] <- raster::stack(mod_preds, qminmax, qrange)
    
  }
  
  # model_stack[sapply(model_stack,is.null)] <- raster(nrow=12500, 
  #                                                    ncol=7000,
  #                                                    crs="+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +ellps=airy +units=m +no_defs")
  
  # To combine them together need to remove the NULL raster layers (i.e. if a model hasn't worked)
  model_stack <- model_stack[!sapply(model_stack,is.null)]
  
  species_stack[[i]] <- raster::stack(model_stack)
  
  # Output the models that failed too
  error_out[[i]] <- do.call('rbind', errored_models) 
  
}

# which models didn't work
errors <- do.call('rbind', error_out)
errors

# name the list entries
names(species_stack) <- names

```


## Create a moth counts raster


```{r counts, echo=FALSE}

##  create a 'counts' raster layer  
dfm <- read.csv('Data/species_data/moth/DayFlyingMoths_East_Norths.csv')
head(dfm)

xy <- dfm[,c("lon","lat")]
spdf.moth <- SpatialPointsDataFrame(coords = xy, data = dfm,
                                    proj4string = CRS("+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +ellps=airy +units=m +no_defs")) 

# plot(spdf.moth) ## check with plot

### create temporary raster to store number of records in each cell ###
moth_counts <- species_stack[[1]][[1]] # get a single raster

# make a raster of zeroes for input
moth_counts[!is.na(moth_counts)] <- 0

# get the cell index for each point and make a table:
counts = table(cellFromXY(moth_counts,spdf.moth))

# fill in the raster with the counts from the cell index:
moth_counts[as.numeric(names(counts))] <- counts
plot(moth_counts)
hist(moth_counts) ## loads of cells with no counts in so probably useless as a layer

```


loads of cells with no counts in so probably useless as a layer


## kernel density


```{r kern_dens, fig.height=8, fig.width=8, message=FALSE, warning=FALSE}

sigma = 10000 # decide number of metres to calculate points over

moth_ppp <- ppp(x = dfm[,c("lon")], y = dfm[,c("lat")],
                owin(xrange = c(extent(moth_counts)[1], extent(moth_counts)[2]),
                     yrange = c(extent(moth_counts)[3], extent(moth_counts)[4])))


dp <- density(moth_ppp, sigma = sigma) 
dpr <- raster(dp)
re_dpr <- resample(dpr, moth_counts, method = 'ngb')
kernel_moth <- mask(re_dpr, moth_counts) # mask to remove sea
plot(kernel_moth)

kern_dens <- function(data, sigma, rast){
  
  
  moth_ppp <- ppp(x = data[,c("lon")], y = data[,c("lat")],
                  owin(xrange = c(extent(rast)[1], extent(rast)[2]),
                       yrange = c(extent(rast)[3], extent(rast)[4])))
  
  
  dp <- density(moth_ppp, sigma = sigma) 
  dpr <- raster(dp)
  re_dpr <- resample(dpr, rast, method = 'ngb')
  kernel_moth <- mask(re_dpr, rast) # mask to remove sea
  
}

par(mfrow = c(2,2))

for(i in c(1000, 5000, 10000, 20000)){
  
  out <- kern_dens(data = dfm, rast = moth_counts, sigma = i)
  
  plot(out, main = paste('grouping distance =', i))
  
}


for(i in c(1000, 5000, 10000, 20000)){
  
  out <- kern_dens(data = dfm, rast = moth_counts, sigma = i)
  
  out_inv <- 1/out
  
  plot(log(out_inv), main = paste('inversed grouping distance =', i))
  
}

par(mfrow = c(1,1))

```

## crop species to smaller scale


```{r cropping, warning=FALSE}

# set location 
location = c(-2.730696, 54.026759) # quernmore
# location = c(-1.110557, 51.602436) # wallingford

# distances
distance = 5000


registerDoParallel(7)

# out_cropped <- list()
system.time(
  out_cropped <- foreach(s = 1:length(species_stack)) %dopar% {
    
    print(s)
    
    sp <- species_stack[[s]]
    
    
    # crop the prediction
    crop_pred <- filter_distance(obj = subset(sp, grep(pattern = 'mean_pred',
                                                       names(sp))),
                                 method = 'buffer',
                                 distance = distance,
                                 location = location)
    
    # crop the error
    crop_err <- filter_distance(obj = subset(sp, grep(pattern = 'quantile_range',
                                                      names(sp))),
                                method = 'buffer',
                                distance = distance,
                                location = location)
    
    if(length(names(crop_pred))>1){
      # get the mean
      m_pred_av <- calc(crop_pred,
                        mean, na.rm = T)
      names(m_pred_av) <- 'predictions'
      
      
      m_quant_av <- calc(crop_err,
                         mean, na.rm = T)
      names(m_quant_av) <- 'error'
    } else {
      
      m_pred_av <- crop_err
      m_quant_av <- crop_err
      
    }
    
    out_rasts <- list(m_pred_av, m_quant_av)
    names(out_rasts) <- c('predictions', 'quantile_var')
    
    return(out_rasts)
    
  }
)

registerDoSEQ()

names(out_cropped) <- names

```


```{r crop_moth_count}

kern_dens(data = dfm, rast = moth_counts, sigma = 5000)


```



## different DECIDE score options


```{r score, fig.height=8, fig.width=10}

score <- recommend_metric(prediction_raster = out_cropped$Adscita_geryon$predictions,
                          error_raster = out_cropped$Adscita_geryon$quantile_var)

par(mfrow = c(2,2))
plot(out_cropped$Adscita_geryon$predictions, main = 'prediction')
plot(out_cropped$Adscita_geryon$quantile_var, main = 'variation')

plot(score$multiply, main = names(score)[1])
plot(score$additive, main = names(score)[2])
par(mfrow = c(1,1))


```





Join the different species together by averaging the decide score weighted by the mean probability of presence across the subsetted area to remove species with low prob of being there... Or maybe weight by the maximum probability of presence so that species only present in a very specific habitat aren't too downweighted 








